Ground state energy of quantum dots using the coupled cluster method

### 2018-04-26 project_2/article/article.tex
Start on appendix for diagrams
Set mass to unity?
de-excitation?

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + W
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $W$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or de-exciting the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

@@

        \subsubsection{Coupled cluster doubles energy equation}
            Using the doubles approximation with the cluster operator $T_2$
            defined in \autoref{eq:T_2} the energy equation becomes
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat e^{-T_2}H_N e^{T_2}\kslat.
            \end{align}
            As the doubles cluster operator doubly excites the reference and
            using the expansion in \autoref{eq:sim_norm_hamiltionian_expansion}
            we see that we can write the energy equation as
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat H_N \kslat + \bslat (H_N T_2)_c\kslat,
            \end{align}
            as the Hamiltonian is only able to de-excite one pair of single
            particle functions. By construction we have that
            \begin{align}
                \bslat H_N \kslat = 0.
            \end{align}
            In the second term only the normal ordered two-body operator can
            contribute as the cluster operator gives a total excitation of $+2$.
            As we are projecting onto the reference we have to de-excite to zero
            again. The normal ordered Fock operator is at most able to excite
            and de-excite by $1$ and does therefore not contribute to the
            overall expression.
            \begin{align}
                \bslat (W_N T_2)_c \kslat
                &= \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab}.
            \end{align}
            In total the energy equation reduces to
            \begin{align}
                \eccd
                &= \sum_{i} h_i^i + \half \sum_{ij}\bra{ij}\ket{ij}
                + \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab},
            \end{align}
            where the first two terms come from the reference energy as shown in
            \autoref{eq:reference_energy}.

@@

>>>>

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    % Why not write $\hbar = c = e = m = 1$?
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + W
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $W$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or relaxing the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

@@

        \subsubsection{Coupled cluster doubles energy equation}
            Using the doubles approximation with the cluster operator $T_2$
            defined in \autoref{eq:T_2} the energy equation becomes
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat e^{-T_2}H_N e^{T_2}\kslat.
            \end{align}
            As the doubles cluster operator doubly excites the reference and
            using the expansion in \autoref{eq:sim_norm_hamiltionian_expansion}
            we see that we can write the energy equation as
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat H_N \kslat + \bslat (H_N T_2)_c\kslat,
            \end{align}
            as the Hamiltonian is only able to relax one pair of single
            particle functions. By construction we have that
            \begin{align}
                \bslat H_N \kslat = 0.
            \end{align}
            In the second term only the normal ordered two-body operator can
            contribute as the cluster operator gives a total excitation of $+2$.
            As we are projecting onto the reference we have to relax to zero
            again. The normal ordered Fock operator is at most able to excite
            and relax by $1$ and does therefore not contribute to the
            overall expression.
            \begin{align}
                \bslat (W_N T_2)_c \kslat
                &= \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab}.
            \end{align}
            In total the energy equation reduces to
            \begin{align}
                \eccd
                &= \sum_{i} h_i^i + \half \sum_{ij}\bra{ij}\ket{ij}
                + \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab},
            \end{align}
            where the first two terms come from the reference energy as shown in
            \autoref{eq:reference_energy}.

@@

\section{Coupled cluster doubles diagrams}
    In order to get an expression for the energy equation and the amplitude
    equations we use a diagrammatic approach.
